\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-01-20 redaction/dhp_sept2016.tex
Changes_redac_2001_bis

The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who point that the overall transport cost for the United States declined from 6 percent to 4 percent between 1974 and 1996. For the same years, our own results point to a total decrease from 6.9 to 4.2\% on average for air transport, and from 9.8\% to 4.8\% for ocean transport.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (say, until 1980) for Ocean transport. \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where quality XXX} Once again, the fact that our estimates are on average higher than those by \citet{Hummels_1999} should not be a surprise, since ours are simple averages, whereas \citet{Hummels_1999}'s are trade-weighted. But the fact that trends are very close are comforting.
%\todo{This has to be clarified: What data uses Hummels? Does he distinguish at some point between air and vessel? }

It appears therefore that the magnitude of transport costs is higher for ocean shipping than for air transport. This result holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of additive costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period. Second, in both transport modes, the additive component appears of sizeable importance As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs. We come back to this aspect in Section \ref{sec:results_trends}.}

\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}

% JH: XXX We have to discuss this: "Overall, these estimates are in line with the ones by \citet{Hummels_1999}, who finds on the same date that the all-commodities trade-weighted average transport cost is 3.8 percent of the fob price for the USA in 1994. More precisely, for this specific year, we find... something slightly higher...

%How do our results compare with other studies? Here, compare with Irrarazabal, Hummels and co.

>>>>

The overall trend pictured in these estimates is in line with the findings by \citet{Hummels_1999}, who point that the overall transport cost for the United States declined from 6 percent to 4 percent between 1974 and 1996. For the same years, our own results point to a total decrease from 6.9 to 4.2\% on average for air transport, and from 9.8\% to 4.8\% for ocean transport.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (say, until 1980) for Ocean transport. \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on freight rates.} Once again, the fact that our estimates are on average higher than those by \citet{Hummels_1999} should not be a surprise, since ours are simple averages, whereas \citet{Hummels_1999}'s are trade-weighted. But the fact that trends are very close are comforting. It appears therefore that the magnitude of transport costs is higher for ocean shipping than for air transport.

Besides, we go beyond by distinguishing within these evolutions between multiplicative and additive transport costs. First, the decreasing trend holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of additive costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period. Second, in both transport modes, the additive component appears of sizeable importance. As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs. We come back to this aspect in Section \ref{sec:results_trends}.}

Before, we check how our estimates of additive costs compare to those of \citet{Irrazabal_2015}, which provide the most recent evidence on that matter. Based on Norwegian firm-level trade data for 2004, \citet{Irrazabal_2015} find that additive costs are on average 14\% of the median price. Expressed in our own terminology, this means they find that $\frac{t/\tau}{\widetilde{p}}=14\%$. Let us define the ratio of additive costs over total costs as $\frac{t}{\tau\widetilde{p}+t}$, and divide both numerators and denominators by $\tau\widetilde{p}$. We get: $\frac{t}{\tau\widetilde{p}+t}=\frac{t/\tau\widetilde{p}}{1+t/\tau\widetilde{p}}$. Plugging \citet{Irrazabal_2015} result gives 0,14/1+0,14= 12,5\%. Additive trade costs represents 12.5\% of total trade costs. Based on \citet{anderson_wincoop_jel}, we know that transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs. Given that in 2004, our estimates for additive (multiplicative) transport costs is 2.9\% (2.7\%) for Vessel, the same reasoning as before leads to 2.9/(2.7+2.9)*11/74 = 7.7\% of total trade costs, that is 62\% (7.7/12.5=62\%) of \citet{Irrazabal_2015}'s estimates. For Air transportation, our estimates for additive (multiplicative) transport costs is 1.5\% (1.9\%), and the same computation brings that additive costs represent 6.6\% of total trade costs, that is 53\% of \citet{Irrazabal_2015}'s estimates. In other words, we find that additive transport costs represent between half and two thirds of the total additive trade costs, which seems totally consistent with the magnitudes reported by e.g. \citet{anderson_wincoop_jel}.

%\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}
